Why robot programming is hard

### 2013-07-18 paper.tex
Another checkpoint. Working through the challenges sections.

Both demonstration and teleop suffer the limitation that a human operator must be immediately in the loop while the robot is operating. To make robots useful in the home, they will need to function {\em autonomously}, which leads to the notion of robot programming: giving high-level directives to robots so that they can function on their own without direct human guidance.  A full programming solution will likely involve a blend of different techniques including demonstration, teleop, and autonomous subroutines. For example, a task of setting the dinner table may include a high-level program of fetching silverware from the kitchen and placing it on the dining table, a demonstration of how to open this particular style of cabinet, and teleoperation when something goes awry (such as the silverware drawer being empty).

@@

Yet our experience with Robots for Humanity underscores how difficult it is for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door could take Henry upwards of 15 minutes to accomplish.  While Henry has only limited interaction capabilities due to motor impairment, he is also strongly motivated. Ordinary able-bodied consumers, on the other hand, would quickly lose patience. If basic robot control via teleoperation is so time-consuming, what hope do we have that people will ever get around to the general task of robot programming?

Granted, designing programming interfaces for motor impaired users is like playing the piano with both hands tied behind your back. Yet while Henry's case highlights the challenges of creating usable interfaces for assistive robotics, it also showcases the enormous benefit to humanity of enabling disabled adults to regain some of their independence.

Until users are able to instruct robots quickly and with low effort, personal robotics will be limited to niche markets.  Why is end user robot programming so challenging? Several compounding factors make it more difficult than normal computer use.

@@

ROS provides a distributed publish/subscribe messaging platform for a collection of {\em nodes} to communicate with each other. Each node encapsulates a small amount of robot functionality, and interfaces with the rest of the ROS ecosystem to make that functionality available through a common interface. For example, ROS nodes exist to stream video from webcams, control the motors in robot arms, plan a path from one point to another, and recognize graspable objects in an image.

@@

\begin{figure}[tbh]
\scalebox{0.15}{\includegraphics{rviz.png}}
\caption{RViz interface for controlling robots}
\label{rviz}
\end{figure}

@@

This interface, while well suited for roboticists who need to examine the low-level sensor data and messages being processed by the robot, is not very user-friendly. It also illustrates some of the difficulties inherent in robot programming, described below.

\section{High dimensionality}

@@

Unlike traditional computer interfaces which are restricted to two dimensions, robot control requires manipulating objects in a three-dimensional world. Visualizing a 3D world on existing 2D displays requires the extra dimension to be projected down into two dimensions, giving rise to the popular ``first person shooter'' style of interface. These interfaces display the world as perceived from a virtual camera, where objects can be occluded by other objects. Additional camera controls are needed to pan/zoom around the space to overcome occlusion, and these controls can be nonintuitive to non-expert computer users.

High dimensionality is even more important in specifying how to grasp and place objects in the world.  In many cases, the robot itself might occlude the object it is trying to manipulate. For example, if a robot uses its arm to pick up an object on a table in front of it, the arm might get in the way of a sensor mounted on its head, preventing the robot from seeing the exact thing it is trying to grasp! One solution developed by roboticists has been to include additional cameras on different parts of the robot, such as a forearm-mounted camera, to get a better view on the object being picked up. However, making sense of multiple translated and rotated video streams can be confusing for those of us accustomed to seeing only from our own head-mounted eye sensors!

@@

% TODO: introduce DOF acronym earlier

>>>>

Both demonstration and teleop suffer the limitation that a human operator must be immediately in the loop while the robot is operating. To make robots useful in the home, they will need to function {\em autonomously}, which leads to the notion of robot programming: giving high-level directives to robots so that they can function on their own without direct human guidance.  A full programming solution will likely involve a blend of different techniques including demonstration, teleop, and autonomous subroutines. For example, a task of setting the dinner table may include a high-level program of fetching silverware from the kitchen and placing it on the dining table, a demonstration of how to open this particular silverware drawer, and teleoperation when something goes awry (such as the silverware drawer being empty).

@@

Yet our experience with Robots for Humanity underscores how difficult it is for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door could take Henry upwards of 15 minutes to accomplish.  While Henry has only limited interaction capabilities, he is also strongly motivated because he has no alternatives. Able-bodied consumers, on the other hand, would quickly lose patience. If even basic robot control via teleoperation is so time-consuming, what hope do we have that people will ever get around to the general task of robot programming?

Granted, designing teleoperation interfaces for motor-impaired users is deliberately trying to solve the hardest problem first.  Yet while Henry's case highlights the challenges of creating usable assistive robotics interfaces, it also illustrates the enormous benefit to humanity of enabling disabled adults to regain some of their independence.

With this goal in mind, autonomous programmable robots are more than a consumer toy, but a technology that could changes people's lives for the better. Why, then, does robot programming for the masses remain such a difficult challenge? What makes robots so difficult to control?

The explanation begins with a look at the state of the art in robot software today.

% Until users are able to instruct robots quickly and with low effort, personal robotics will be limited to niche markets.  Why is end user robot programming so challenging? Several compounding factors make it more difficult than normal computer use.

@@

ROS provides a distributed publish/subscribe messaging platform for a collection of {\em nodes} to communicate with each other. Each node encapsulates a small amount of robot functionality, and interfaces with the rest of the ROS ecosystem to make that functionality available through a common interface. For example, ROS nodes exist to stream sensor data from webcams and laser rangefinders, control the motors in robot arms, plan a path from one point to another, and recognize graspable objects in an image.

@@

\begin{figure*}[tbh]
\center\scalebox{0.20}{\includegraphics{rviz.png}}
\caption{RViz interface for controlling robots}
\label{rviz}
\end{figure*}

@@

This interface, while well suited for roboticists who need to examine the low-level sensor data and messages being processed by the robot, is not very user-friendly. It also illustrates some of the difficulties inherent in robot programming.

\section{A 3D world}

@@

Unlike traditional computer interfaces which are limited to two dimensions, robot control requires manipulating objects in a three-dimensional world. Visualizing a 3D world on existing 2D displays requires the extra dimension to be projected down into two dimensions, giving rise to the popular ``first person shooter'' style of interface. These interfaces display the world as perceived from a virtual camera, where objects can be occluded by other objects. Additional camera controls are needed to pan/zoom around the space to overcome occlusion. These controls are often non-intuitive, even to seasoned computer users.

Occlusion presents a challenge because robots must perceive an object in order to manipulate it.  Often times, the robot itself might occlude the object it is trying to manipulate! For example, consider a humanoid robot with head-mounted vision sensors using an arm to pick up an object from the table in front of it. While moving the arm into position to grasp the object, the arm will likely come between the head and the object, obscuring the robot's vision.

One solution has been to mount additional cameras on different parts of the robot, such as on the robot's forearm or wrist, to get a better view on the object being picked up. However, making sense of multiple translated and rotated video streams can be confusing for those of us accustomed to perceiving the world only through head-mounted eye sensors!

Typical input devices such as mice and touchscreens are also only good for interacting in two dimensions. For example, tapping on a point in a camera image does not uniquely identify a point in 3D space, but a line of possible points. When telling a robot to put something down, how can it know which of those points was meant? Current interfaces, such as the RViz display shown in Figure~\ref{rviz}, require complex camera manipulation and precise 3D positioning in order to uniquely specify the desired location.

\section{Arm complexity}

A different aspect of high-dimensionality comes from the design of robot arms. A solid object in 3D space has six degrees of freedom (DOF): three dimensions to specify its position, and three more to specify its rotation. If a robot arm had only 6-DOF and was trying to execute a specific grasp of a particular object, such as picking up soda can from the side, there would be only a single arm configuration that would allow the gripper to pick up the can. In contrast, human arms and advanced robot arms have 7-DOF, which allows them one extra dimension of flexibility in achieving those grasps. To visualize this, imagine grasping a soda can on the table in front of you and then flapping your elbow up and down while keeping the can stationary. That extra degree of freedom enables your arm to take on all those different shapes while still achieving the desired grasp.

Interfaces for robot arm teleoperation typically involve setting the position or angle of each joint in the arm individually. For example, a tablet-style interface might have fourteen buttons, one pair used to increase or decrease the angle of each joint in the arm. That additional degree of freedom makes it easier to identify a valid arm configuration for executing a grasp, since only one of many valid configurations must be found, not the unique one.

For teleoperators who are manually positioning a robot arm must only find one way to position the various joints

It may also make it easier for teleoperators to find a valid arm configuration to reach a desired grasp. Yet those seven degrees of freedom also represent seven variables to control when using a direct-manipulation teleop interface

@@
